---
title: "Capstone"
author: "Rahul"
date: "December 12, 2016"
output: html_document

*Springboard Foundations of Data Science*  
*By Rahul Arora*  
*Mentor: Shmuel Naaman*  
*December, 12, 2016*

## Overview
Currently, daily inventory calculations are performed by direct delivery sales employees who must single-handedly predict the supply and demand based on their personal experiences with each store. With some bakery items carrying a one week shelf life, the acceptable margin for error is small. Forecasting inventory demand will increase customer satisfaction and reduce surplus product unfit for sale.

## Data set
Downloaded data from Kaggle (https://www.kaggle.com/c/grupo-bimbo-inventory-demand/data)
Dataset consists of 9 weeks of sales transactions in Mexico. Every week, there are delivery trucks that deliver products to the vendors. Each transaction consists of sales and returns. Returns are the products that are unsold and expired. The demand for a product in a certain week is defined as the sales this week subtracted by the return next week.



## Load Libraries

```{r library_load echo=FALSE, message=FALSE, warning=FALSE}

library(caret)
library(randomForest)
library(ggplot2)
library(gridExtra)
library(dplyr)

df_train_random <- read.csv('C:\\Users\\Rahul\\Documents\\SpingBoard\\Capstone\\Bimbo\\train_random.csv')

```


## Data Analysis
Explore the various features of the Data set

```{r plot_features echo=FALSE, message=FALSE, warning=FALSE}

summary(df_train_random)

# Adjusted Demand 

ggplot(data = df_train_random, aes( x = Adjusted_Demand ) ) +  
  geom_histogram( bins = 30 ) +scale_x_log10()

#  Used Box Plot to observe outliers in Adjusted Demand

ggplot(data = df_train, aes( x = 1, y = Sales_this_week.pesos. ) ) +  
  geom_boxplot()   +scale_y_log10()

# Product ID

df_train_random %>% count(Product_ID, sort = TRUE) -> product_count
top10_products = product_count$Product_ID[1:10]

df_train_random %>% group_by(Product_ID) %>%
  summarise( Sum_Adj_Demand = sum(Adjusted_Demand),
             Sum_Sales_Unit = sum(Sales_unit_this_week),
             sum_Returns_Unit = sum(Returns_unit_next_week)) -> Products_summary

View(Products_summary)

Products_summary %>% filter(sum_Returns_Unit ==0) -> High_demand_Products

Products_summary %>% filter(Sum_Sales_Unit ==0) -> Low_demand_Products

View(High_demand_Products)

View(Low_demand_Products)

df_train_random %>% 
    filter(Product_ID %in% top5_products) %>% 
    ggplot(aes(x = Product_ID, y = log1p(Adjusted_Demand))) +
    geom_boxplot() +
    ggtitle("Boxplots of Demand per Product")

df_train_random %>% 
    filter(Product_ID %in% High_demand_Products$Product_ID) %>% 
    filter(Adjusted_Demand < 50) %>% 
    ggplot(aes(x = Adjusted_Demand, fill = Product_ID)) +
    geom_bar() +
    facet_wrap( ~ Product_ID) +
    ggtitle("Demand Distributions for Top Products")
    
df_train_random %>% 
    filter(Product_ID %in% Low_demand_Products$Product_ID) %>% 
    filter(Adjusted_Demand < 50) %>% 
    ggplot(aes(x = Returns_unit_next_week)) +
    geom_bar() +
    facet_wrap( ~ Product_ID) +
    ggtitle("Demand Distributions for Bottom Products")
    
# Client ID 

df_train_random %>% count(Client_ID, sort = TRUE) -> client_count
top10_client = client_count$Client_ID[1:10]

client_count %>% filter(Client_ID %in% top5_client) -> Top10_clients

ggplot(data = Top10_clients, aes( x = factor(Client_ID), y = n ) ) +
geom_bar(stat = "identity") +
 scale_x_discrete(name="Client ID")+
  scale_y_log10(name="Counts") +
ggtitle("Top Clients")


# Sales Channel

df_train_random %>% count(Sales_Channel_ID, sort = TRUE) -> Sales_Channel_count

ggplot(data = Sales_Channel_count, aes( x = factor(Sales_Channel_ID), y = n ) ) +
geom_bar(stat = "identity") +
 scale_x_discrete(name="Sales_Channel ID")+
  scale_y_log10(name="Counts") +
ggtitle("Top Sales Channel")
```

## Feature Selection

```{r feature echo=FALSE, message=FALSE, warning=FALSE}

inTraining <- createDataPartition(df_train_random$Adjusted_Demand, p = .8, list = FALSE)

# Split the Training data set into Train and Test data frames

train_1 <- df_train_random[ inTraining,]
test_1  <- df_train_random[-inTraining,]

A<- summary(lm(Adjusted_Demand ~., data = df_train_random))

# fit the model
model <- train(Adjusted_Demand ~., 
               data=train_1, 
               method="rf", 
               metric="RMSE", 
               tuneGrid=expand.grid(.mtry=4), 
               ntree=400,
               importance=TRUE)

# what are the important variables (via permutation)
vi <- varImp(model, type=1)
plot(vi, top=10, main = 'Feature Importance')


```


## Random Forest Model: K-Fold and Grid Search Cross Validation 

In k-fold cross-validation, the original sample is randomly partitioned into 'k' equal sized subsample. 
Of the 'k' subsamples, a single subsample is retained as the validation data for testing the model, and the remaining 'k−1' subsamples are used as training data. 
The cross-validation process is then repeated 'k' times (the folds), with each of the 'k' subsamples used exactly once as the validation data. The 'k'results from the folds can then be averaged to produce a single estimation.

```{r model echo=FALSE, message=FALSE, warning=FALSE}

# ensure results are repeatable
set.seed(123)

# Manual Grid Search
control <- trainControl(method="repeatedcv", 
                        number=10, 
                        repeats=1, 
                        search="grid")
tunegrid <- expand.grid(.mtry=c(3,4))
modellist <- list()
metric = "RMSE"

#For loop of gridsearch
for (ntree in c(50, 100, 150, 200, 250, 300, 350, 400, 450, 500)) {
	set.seed(123)
	fit <- train(Adjusted_Demand ~ Product_ID + Sales_unit_this_week + Sales_this_week.pesos. +
                 Returns_unit_next_week + Client_ID + Route_ID + 
                 Sales_Channel_ID + Returns_next_week.pesos., 
	             data=train_1, 
	             method="rf", 
	             metric=metric, 
	             tuneGrid=tunegrid, 
	             trControl=control, 
	             ntree=ntree)
	key <- toString(ntree)
	modellist[[key]] <- fit
}

results <- resamples(modellist)
summary(results)

```


## Random Forest Model
***
Random forests are an ensemble learning method for regression, that operate by constructing a multitude of decision trees at training time and outputting the mean prediction (regression) of the individual trees. 
Each decision tree is constructed by using a random subset of the training data. After you have trained your forest, you can then pass each test row through it, in order to output a prediction.
 
```{r RF echo=FALSE, message=FALSE, warning=FALSE}

# fit the randomforest model
model_rf <- randomForest(Adjusted_Demand ~ Product_ID + Sales_unit_this_week + Sales_this_week.pesos. +
                 Returns_unit_next_week + Client_ID + Route_ID + 
                 Sales_Channel_ID + Returns_next_week.pesos., data=train_1, mtry=3, ntree=200)

# predict the outcome of the training data
predicted_tr <- predict(model_rf, newdata=train_1, select = -c(Adjusted_Demand))
actual_tr <- train_1$Adjusted_Demand
rsq_tr <- 1-sum((actual_tr-predicted_tr)^2)/sum((actual_tr-mean(actual_tr))^2)

# predict the outcome of the testing data
predicted <- predict(model_rf, newdata=test_1, select = -c(Adjusted_Demand))
actual <- test_1$Adjusted_Demand
rsq <- 1-sum((actual-predicted)^2)/sum((actual-mean(actual))^2)

```

## Evaluating the Model with Validation Data, unseen by the model
The evaluation metric for this competition is Root Mean Squared Logarithmic Error.

```{r Eval echo=FALSE, message=FALSE, warning=FALSE}

df_test_final <- read.csv('C:\\Users\\Rahul\\Documents\\SpingBoard\\Capstone\\Bimbo\\train_week9_random_subset.csv')

# predict the outcome of the Validation data (Data unseen by the model)
predicted_eval <- predict(model_rf, newdata=df_test_final, select = -c(Adjusted_Demand))
actual_eval <- df_test_final$Adjusted_Demand
rsq_eval <- 1-sum((actual_eval-predicted_eval)^2)/sum((actual_eval-mean(actual_eval))^2)

## Model Performance - Root 
rmlse <- function(model_rf) { 
  y <- train_1$Adjusted_Demand
  y.pred <- predict(model_rf, df_test_final)
  return(sqrt(1/length(y)*sum((log(y.pred +1)-log(train_1$Adjusted_Demand +1))^2)))
}

rmlse(model_rf)

```

